---
title: "Homework 13"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Unit 13 Assignment

###Chapter 20, Problem 7
Use the now-familiar lowbwt data set.

a. Using germinal matrix hemorrhage as the response, fit a logistic model where x1 is five minute Apgar score. Write the equation and interpret β1, the estimated coefficient of Apgar score.
```{r}
library(readr)
lowbwt <- read_csv("~/Papers/Biostatistics JHU 2021/lowbwt.csv")

fit1 <- glm(grmhem ~ apgar5, data = lowbwt, family = binomial)
fit1
exp(-0.2496)
```
b. If a particular child has a five-minute Apgar score of 3, what is the predicted probability that this child will experience a brain hemorrhage? What is the probability if the child's score is 7?
```{r}
logodds <- -.3037 - .2496*3
odds <- exp(logodds)
p <- odds / (1+odds)
p
```

c. At the 0.05 level of significance, test the null hypothesis that the population parameter β1 is equal to 0. What do you conclude?
*reject the null hypothesis because it is not 0.*

d. Now fit the regression model where x2 represents toxemia status. Interpret β2, the estimated coefficient of toxemia.
```{r}
fit2 <- glm(grmhem ~ tox, data = lowbwt, family = binomial)
fit2
```

e. For a child whose mother was diagnosed with toxemia during pregnancy, what is the predicted probability of experiencing a germinal matrix hemorrhage? What is the probability for a child who was not diagnosed with toxemia?
```{r}
logodds <-  -1.535 - 1.460
odds <- exp(logodds)
p2 <- odds / (1+odds)
p2
logodds <-  -1.535 - 1.460
odds <- exp(logodds)
p3 <- odds / (1-odds)
p3
```
f. What are the estimated odds of suffering a germinal matrix hemorrhage for children whose mothers were diagnosed with toxemia relative to children whose mothers are not.
```{r}
logodds <-  -1.535 - 1.460
odds <- exp(logodds)
p4 <- odds / (1+odds)
p4
```

g. Construct 95% confidence interval for the population odds ratio. Does this interval contain the value 1? What does this tell you?
```{r}
t.test(lowbwt)
```
*The CI is between 12.34144 15.33189 and does not include 1, so this tells me there is not a significant difference between children with mothers who suffer from toxemia and those whose mothers do not.*

###Chapter 21, Problem 7
In the table below you will find survival times in months since diagnosis for 10 AIDS patients suffering from concomitant esophageal candidiasis, an infection due to Candida yeast, and cytomegalovirus, a herpes infection that can cause serious illness. Censored observations are denoted by a plus sign. 

Patient Number |	Survival (months)
-----------|-----------
1	| 0.5+
2	| 1
3	| 1
4	| 1
5	| 2
6	| 5+
7	| 8+
8	| 9
9	| 10+
10	| 12+

a. How many deaths were observed in this sample of patients?
*There were 4 deaths in this sample set.*

b. Estimate the survival function S(t). 
*10/11, 7/10, 6/9, 5/8, 4/7, 3/6, 2/5, 1/4*

c. What is S(1), the estimated probability of survival at one month? What is the estimated probability at 5 months? At 6 months?
*S(1)=0.7, S(5)=.625, S(6)=.625*

d. Construct a survival curve.

###Chapter 21, Problem 9
In a study of bladder cancer, tumors were removed from the bladders of 86 patients. Subsequently, the individuals were assigned to be treated with either a placebo or the drug thiopeta. Time to recurrence of tumor in months is saved under the variable name time in the data set *bladder*. Treatment status is saved under the name group; the 1 value represents placebo. Indicators of censoring status; where 0 indicates a tumor did not occur and the observation was censored, is saved under the name censor.

a. Use a Kaplan-Meier curve to estimate the survival function in each treatment group, and plot.
```{r}
library(readr)
bladder <- read_table2("~/Papers/Biostatistics JHU 2021/bladder.txt")
t.test(bladder$time, var.equal = TRUE)
Surv(time=bladder$time, event=bladder$censor)
```

b. Does it appear that the individuals in one group have a longer time to first recurrence than in the other group?
*It does appear that one group, thiopta, has the longer interval of recurrence of tumors.* 

c. Test the null hypothesis that the distributions of recurrence times are identical. What do you conclude?
*The null hypothesis can be rejected as the p-value is close to 0.*


