---
title: "Homework 3"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Assignment for Biostatistics                                                         Week 3:

Read in the cholesterol data file from the class webpage. 
```{r}
#Enter code here

library(readr)
cholesterol_1_ <- read_csv("~/Papers/Biostatistics JHU 2021/cholesterol(1).txt")


```


1.    Make a histogram of the total cholesterol readings in this dataset, with the y-axis equal to the probability of being in each bar (not the frequency). Estimate the mean and standard deviation of this normal distribution.  Give the estimates of these parameters, super-impose a normal distribution with that mean and SD on the histogram. 
```{r}
#Enter code here
hist(cholesterol_1_$TOTCHOL, probability = T)
x <- seq(60,290,.01)
y <- dnorm(x,190.925,35.52988)
lines(x,y,col= "red",lw=3)



```


2. Create a new variable that is a standardized version of total cholesterol by subtracting off the mean and dividing by the standard deviation.  Make another histogram, this time of the standardized data. Superimpose a standard normal curve (“standard normal”: mean=0, SD=1)
```{r}
#Enter code here

hist(cholesterol_1_$TOTCHOL, probability = T)
x <- seq(y-5.374702)
y2 <- dnorm(x,.01)
lines(x,y2, col="red", lwd=2)



```


3. The American Heart Association (previously) defined “High (Total) Cholesterol” as 240 or above.  What proportion of the people in this dataset have total cholesterol of 240 or above? 

```{r}
#Enter code here
above240<- (cholesterol_1_$TOTCHOL>=240)
sum(above240)
length(cholesterol_1_$TOTCHOL)

```

4. Imagine you wanted to compare the population from which this sample is drawn to a different population, in which it is well-documented that 10% of people have total cholesterol of 240 or above.    

```{r}
#Enter code here
```

If the true rate of high cholesterol in this population were 10%, what is the probability that you would see exactly as many people with high cholesterol as you saw here? 

```{r}
#Enter code here
dbinom(33,453,0.10)


```


If the true rate of high cholesterol in this population were 10%, what is the probability that you would see at least as many people with high cholesterol as you saw here? 

```{r}
#Enter code here
1-pbinom(32,453,0.10)


```


5. Read the document on this week’s lesson called “Understanding Simulations in R”.

Conduct a small simulation.  Using 10% as the true rate of high cholesterol, generate data for 453 simulated people; do this 1000 times.  Make a histogram of your simulated counts (note: with 1000 observation, could set nclass to 20 or so).  Where is the count you actually observed on the histogram? Mark it with a vertical line. Does this count seem unusual among samples of people where 10% is the truth? 

```{r}
#Enter code here
rbinom(1000,453,0.10)
simdata<-rbinom(1000,453,0.10)
sum(simdata>=33)/1000
hist(simdata)
abline(v=33,col="red", lwd=2)


```

____________________________________________

## Some commands you might find useful
```
chol <- read.csv("C:/Desktop/Data/cholesterol.txt")
hist(chol$TOTCHOL, probability =T)
mean(chol$TOTCHOL)
sd(chol$TOTCHOL)

#Add a normal curve to a plot
x <- seq( min(chol$TOTCHOL), max(chol$TOTCHOL))
curve(dnorm(x, mean = __ , sd = __ ), add = TRUE, col="red")	

# Gives the binomial density at this point-- i.e. the probability of exactly 
# "x" successes out of "size" trials with "prob" chance of success on each.
dbinom(x= , size= , prob=)
	#Gives the binomial distribution function at this point-- i.e. the probability 
#of "x" or fewer successes out of "size" trials with "prob" chance of success on each.
pbinom(q= , size= , prob=)
#Generates n random binomial observations, each the count of successes out of "size" trials with "prob" chance of success on each.
rbinom(n= , size= , prob=)	

abline(v= , h= ) #Adds a vertical line at v or a horizontal line at h 

```

# Other homework: Chapter 7 #10 #13 & # 20  from Principles of Biostatistics

*I recommend doing the calculations in this homework by hand. Feel free to use R to check yourself, but please work through the formulas a couple of times.* 

10. Consider a group of seven individuals selected form the population of 65- to 74-year olds in the United States. The number of persons in this sample who suffer from diabetes is a binomial random variable with parameters n=7 and p=0.125

a. If you wish to make a list of the seven persons chosen, in how many ways can they be ordered?

b. Without regard to order, in how many ways can you select 4 individuals from this group of 7? 

c. what is the probability that exactly two of the individuals in your sample suffer from diabetes? 

d. what is the probability that four of them have diabetes?

12. According to the Behavioral Risk Factor Surveillance System, 58% of all Americans adhere to a sedentary lifestyle.

a. If you selected repeated samples of size 12 from the US population, what would be the mean number of individuals per sample who do not exercise regularly? What would be the standard deviation?  

b. Suppose you select a sample of 12 individuals and find that 10 of them do not exercise regularly. Assuming the Surveillance System is correct, what is the probability you would have obtained results as bad or worse than those you observed? 


20. In the Framingham study, serum cholesterol levels were measured for a large number of healthy males. The population was followed for 16 years. At the end of the time, the men were divided into 2 groups: those who had developed coronary disease and those who had not. The distributions of the initial serum cholesterol levels for each group were found to be approximately normal. Among individuals who eventually developed coronary disease, the mean serum cholesterol level was μd= 244 mg/100 ml and the standard deviation was σd=51 mg/100ml; for those who did not go on to develop the disease, the mean serum cholesterol was μnd= 219 mg/100 ml and the standard deviation was σd=41 mg/100ml.

a. Suppose that an initial serum cholesterol level of 260 mg/100ml was used to predict coronary heart disease. What is the probability of correctly predicting heart disease for a man who will go on to develop it? 

b. Using the cutpoint in (a), What is the probability of predicting heart disease for a man who will not go on to develop it? 

c. Using the cutpoint in (a), What is the probability of failing to predict heart disease for a man who will develop it? 

d. What would happen to the probabilities of false positives and false negative errors if the cutoff point for predicting heart disease was lowered to 250 mg/100 ml?

e. In ths population, does initial serum cholesterol appear to be useful for predicting coronary heart disease? Why or why not? 
