---
title: "Homework 6"
output:
  word_document: default
  html_document: default
---
#Biostatistics Assignment for Biostatistics , Week 6

Download and open the “Blood Over Time” data from the class webpage. Answer the following questions referring to this dataset.

Same set-up as last week:
A scientist was worried that the quality of the blood they were using for ongoing experiments was degrading with longer storage times. They obtained blood from 40 HIV + volunteers and divided each blood sample up into four parts. They then measured the percent of CD4 and CD8 cells that were expressing CD127 from each sample at 4 hours, 24 hours, 48 hours and 72 hours. 

For the following, we will only use 4 and 72 hour CD8 data. 

1.	Last time, we tested for a difference between 4 and 72 hour CD8 data by creating a new variable that summarized the difference for each person. You can do the same test using the “paired=T” choice, without creating this variable.  Compare the 4 hour and 72 hour CD8 using a paired design.

```{r}

 library(readr)
blood_CD8 <- read_csv("~/Papers/Biostatistics JHU 2021/blood_CD8.csv")
t.test(blood_CD8$`4`, blood_CD8$`72`, paired = T)


```

 
For the following, we will only use 4 hour CD4 data, but will also need to know IL-2 status. 

2.	Some of the individuals in the sample had taken IL-2 in the recent past when their blood was drawn; others had not. Do a t-test to answer the question as to whether there is a difference in the CD4% between people who have and have not taken IL-2, based on the 4 hour data. Write a one-sentence conclusion (including the p-value) of the results of the test, the way you might include it in a science paper. Do the test:
a.	Assuming equal variances.

```{r}
 library(readr)
Blood <- read_csv("~/Papers/Biostatistics JHU 2021/Blood over time(1).csv")

t.test(Blood$CD4.xCD127[Blood$Hours.post.draw==4]~Blood$IL2[Blood$Hours.post.draw==4], var.equal=T)


```




b.	Assuming unequal variances.


```{r}

t.test(Blood$CD4.xCD127[Blood$Hours.post.draw==4]~Blood$IL2[Blood$Hours.post.draw==4])

```




*c.	(FYI only, not due yet) The researcher comes back to you a few weeks later and wants to design two new trials to address the same questions, with higher statistical power. How many subjects to answer the following question: Are there differences between people with/without a history of IL2? Assume that CD4% in both groups will be normally distributed with a standard deviation of 8%. How many people are needed PER GROUP in order to have 80% power to detect a difference of 5% or greater? (on your own)*


```{r}

```


Chapter 11, Problem 5
A crossover study was conducted to investigate whether oat bran cereal helps to lower cholesterol levels in hypercholesterolemic males. Fourteen such individuals were randomly placed on a diet that included either oat bran or corn flakes; after two weeks, their low-density lipoprotein (LDL) cholesterol levels were recorded. Each man was then switched to the alternative diet. After a second two-week period, the LDL cholesterol level of each individual was again recorded. The data from this study are shown below:

| subject	| corn flakes LDL	| oat bran LDL	| Difference |
|--------|----------------|---------------|------------|
| 1	| 4.61 | 3.84	 | | 
| 2	| 6.42	| 5.57 | | 
| 3	| 5.40	| 5.85	  | | 
| 4	| 4.54	| 4.80	  | | 
| 5	| 3.98	| 3.68	  | | 
| 6	| 3.82	| 2.96	  | | 
| 7	| 5.01	| 4.41	  | | 
| 8	| 4.34	| 3.72	  | | 
| 9	| 3.80	| 3.49	  | | 
| 10 | 4.56	| 3.84	  | | 
| 11	| 5.35	| 5.26	  | | 
| 12	| 3.89	| 3.73	  | | 
| 13	| 2.25	| 1.84	  | | 
| 14	| 4.24	| 4.14	  | | 

a.	Are the two samples paired or independent?
*The data is paired because it was conducted on the same 14 people for each cereal.*

b.	What are the appropriate null and alternative hypotheses for a two-sided test
*H0:μd=0 against H1:μd≠0 (two-tailed)*
c.	 Conduct the test at the 0.05 level of significance. What is the p-value?
*The p-value is .0053, so we must reject the null hypothesis.*
d.	What do you conclude?
*Overall, between the 2 tests, the Oat Bran cereal seemed to do better than corn flakes. LDL was lower in the oat bran cereal than it was in the corn flakes cereal.*

##Chapter 11, Problem 8

A study was conducted to determine whether an expectant mother's cigarette smoking has any effect on the bone mineral content of her otherwise healthy child. A sample of 77 newborns whose mothers smoked during pregnancy has mean bone mineral content $\bar x_1 = 0.098 g/cm$ and standard deviation $s_1 = 0.026 g/cm$; a sample of 161 infants whose mothers did not smoke has mean $\bar x_2 = 0.095 g/cm$ and standard deviation $s_2 = 0.025 g/cm$. Assume that the underlying population variances are equal.

a.	Are the two samples paired or independent?
*The two samples are independent. Independent because the mothers who smoked were not specifically matched to mothers who did not smoke during pregnancy.*
b.	State the null and alternative hypotheses of the two-sided test.
*The null and Alternative hypothesis of the two sided test are H0:μ1=μ2 against Ha:μ1≠μ2.*
c.	Conduct the test at the 0.05 level of significance. What do you conclude?
*The p-value is 0.393 which is greater than the significance level of α=0.05, we fail to reject the null hypothesis.*
